\section{Parsing}
\label{sec:Parsing}

The parsing interface of EXIP is similar to SAX and StAX. The main difference is
that most of the XML schema build-in types are passed in a native binary form
rather than as a string representation.
The EXIP parser processes the EXI stream in one direction and produces events
on occurrence of information items such as elements, attributes or content values.
The applications using EXIP must register callback functions for the events that they are interested in.
In this way the data from the EXI stream is delivered to the application as a
parameters to these callback functions.
The header \texttt{contentHandler.h} provides declarations of the callback functions
while the API for controlling the progress through the EXI stream is in \texttt{EXIParser.h}.

\subsection{Schema-less decoding}

When in schema-less mode all value items are encoded with strings and hence
delivered to the application through the \texttt{stringData()} callback handler.
Similarly to the serialization, the parsing consists of 7 simple steps:
\begin{enumerate}
 \item Declare a parser container that holds the serialized data and EXIP state:
\begin{lstlisting}
Parser parser;
\end{lstlisting}

 \item (Optional) Define an external input stream for receiving the EXI document. If an input stream
is not defined the whole EXI document must be stored in a memory buffer before the start of the
parsing process:
\begin{lstlisting}
IOStream in; /* Declare an input stream container */
FILE *exifile; /* Using a file for retrieving the EXI document */
exifile = fopen(sourceEXIFile, "rb" ); /* open the file before use */
in.readWriteToStream = readFromFileInputStream; /* readFromFileInputStream
is a function implementing the actual reading from the stream */
in.stream = exifile; /* Sets the input stream to the file */
\end{lstlisting}

 \item Define an intermediate buffer, application data structure and initialize the parser object:
\begin{lstlisting}
#define INPUT_BUFFER_SIZE 200 /* The size of the buffer in bytes */
char buffer[INPUT_BUFFER_SIZE];

/** The application data that is passed to the callback handlers*/
struct applicationData
{
  unsigned int elementCount;
  unsigned int nestingLevel;
} appData;


initParser(&parser, buffer, INPUT_BUFFER_SIZE, 0, &in, NULL, &appData);

/**
 * @param[in, out] parser EXIP parser container
 * @param[in, out] buffer binary buffer for fetching EXI encoded data
 * @param[in] INPUT_BUFFER_SIZE the size of the buffer in bytes
 * @param[in] 0 the number of bytes that are pre-fetched in the buffer
 * @param[in] in input stream to be used to fill the binary buffer when
		      empty if any; NULL otherwise
 * @param[in] NULL a compiled schema information to be used for schema
		      enabled processing; NULL if no schema is available
 * @param[in] appData application data to be passed to the callback handlers
 */      
\end{lstlisting}

 \item Initialize the application data and register the callback handlers with the parser object:
\begin{lstlisting}
appData.elementCount = 0; /* Example: the number of elements passed */
appData.nestingLevel = 0; /* Example: the nesting level */

parser.handler.fatalError = sample_fatalError;
parser.handler.error = sample_fatalError;
parser.handler.startDocument = sample_startDocument;
parser.handler.endDocument = sample_endDocument;
parser.handler.startElement = sample_startElement;
parser.handler.attribute = sample_attribute;
parser.handler.stringData = sample_stringData;
parser.handler.endElement = sample_endElement;

/** According to the above definitions:
  *  When the parser start parsing the body, the sample_startDocument()
  *  callback will be invoked; when a start of an element is parsed the
  *  sample_startElement() will be invoked and so on. All of the events
  *  for which there is no handler registered will be discarded. 
  */
\end{lstlisting}

 \item Parse the header of the stream:
\begin{lstlisting}
parseHeader(&parser);
/* The header fields are stored in parser.strm.header*/
\end{lstlisting}

 \item Parse the body of the EXI stream, one content item at a time:
\begin{lstlisting}
errorCode error_code = ERR_OK;
while(error_code == ERR_OK)
{
  error_code = parseNext(&parser);
}

/**
 * On successful parsing step, the parseNext() returns ERR_OK if there
 * are more content items left for parsing and PARSING_COMPLETE in case
 * the parsing is complete. If error conditions occur during the
 * process it returns an error code.
 */
\end{lstlisting}

 \item Destroy the parser object and free the memory allocated by it. If any other
streams are left open close them as well:
\begin{lstlisting}
destroyParser(&parser);
fclose(exifile);
\end{lstlisting}

\end{enumerate}
\subsection{Schema-enabled decoding}
When in schema-enabled mode the basic parsing steps are essentially the same.
The difference is in the \texttt{EXIPSchema*} parameter passed to the \texttt{initParser()}.
You need to pass a valid \texttt{EXIPSchema} object and not NULL as in the schema-less
case. This object contains all the definitions and constrains from the XML schema and
its creation is a topic of section \ref{sec:Schema-Infromation} \textbf{Schema Information}.
During parsing, the value types as defined by the schema definitions
are delivered through the corresponding callback handlers and not only \texttt{stringData()}.
For example, \texttt{xsd:integer} is delivered through \texttt{intData()} and
\texttt{xsd:boolean} through \texttt{booleanData()}. Apart from that all other decoding steps
are the same as in schema-less mode.